---
title: "Ejercicios Procesos Estocasticos"
output: html_notebook
---
### Aviso ->
Aca pondre todas las utilidades que muestra el libro en la pagina web para facilitar mucho el uso -> 

Pueden minimizar el código para que no este tan estorbozo.


```{r}
##### Matrix powers ###############################
# matrixpower(mat,k) mat^k
#
matrixpower <- function(mat,k) {
	if (k == 0) return (diag(dim(mat)[1])) 
	if (k == 1) return(mat)
	if (k > 1) return( mat %*% matrixpower(mat, k-1))
 }

###### Simulate discrete-time Markov chain ########################
# Simulates n steps of a Markov chain 
# markov(init,mat,n,states)
# Generates X0, ..., Xn for a Markov chain with initiial
#  distribution init and transition matrix mat
# Labels can be a character vector of states; default is 1, .... k

markov <- function(init,mat,n,labels) { 
	if (missing(labels)) labels <- 1:length(init)
simlist <- numeric(n+1)
states <- 1:length(init)
simlist[1] <- sample(states,1,prob=init)
for (i in 2:(n+1)) 
	{ simlist[i] <- sample(states,1,prob=mat[simlist[i-1],]) }
labels[simlist]
}
####################################################

####################################################
######## Build the transition matrix for random walk on n-cycle
n <- 25
r1 <- c(0,1/2,rep(0,n-3),1/2)
tmat <- matrix(0,nrow=n,ncol=n)
tmat[1,] <- r1 
for (i in 1:(n-1)) tmat[26-i,] <- r1[1 + (i:(n+i-1)) %% n]

### Start the RW on vertex 1
init <- c(1,rep(0,n-1))

sim <- markov(init,tmat,100) # 100 steps
sim
table(sim)/101 # Proportion of visited sites in first 100 steps
sim2 <- markov(init,tmat,100000) #100,000 steps
table(sim2)/100001 # Proportion of visited sites in 100,000 steps
#####################################################

### Stationary distribution of discrete-time Markov chain
###  (uses eigenvectors)
###
stationary <- function(mat) {
x = eigen(t(mat))$vectors[,1]
as.double(x/sum(x))
}

## Example of stationary(mat)
mat <- matrix(c(3/4,1/4,0,1/8,2/3,5/24,0,1/6,5/6),nrow=3, byrow=T)
mat
lambda <- stationary(mat)
lambda
lambda %*% mat

#####################################################
### Standard Brownian motion
### bm(t,n) generates B_0, B_{t/n}, B_{2t/n}, . . . , B_{(n-1)t/n, B_t}
###
bm <- function(t,n) {
	cumsum(c(0,rnorm(n,0,sqrt(t/n))))
}
#### Example of BM plot on [0,15]
tscale <- seq(0,15,15/1000)
plot(tscale,bm(15,1000),type="l")

```





Una cadena de markov tiene esta matriz de transicion ->

 \begin{pmatrix}
0.1 & 0.3 & 0.6\\ 
0 & 0.4 & 0.6\\ 
0.3 & 0.2 &0.5
\end{pmatrix}



$$
P(x_7 = 3 | P_6 = 2)
$$
$$
P(X_9 = 2 | X_1 = 2 , X_ = 1 , X_7 = 3 )
$$
$$
p(x_0 = 3 | x_1 = 1)
$$

$$
E(X_2)
$$


```{r}
matriz <- matrix(c(0.1,0.3,0.6,0,0.4,0.6,0.3,0.2,0.5), ncol = 3, nrow = 3)
matriz
```

a)
```{r}
#Solo localiza 3,2
```
b)
```{r}
matriz_cuadrada <- matriz%*%matriz
matriz_cuadrada
#localizar 3,1
```

c)
$(0.3*0.5) / 0.3*0.4*0.2+0.2*0.3*0.5$

d) 
https://matrixcalc.org/es/#diagonalize(%7B%7B0%2e1,0%2e3,0%2e6%7D,%7B0,0%2e4,0%2e6%7D,%7B0%2e3,0%2e2,0%2e5%7D%7D)


$(0.182,0.273,0.545) * (1,2,3)$ 


2.3 Considerando el modelo de wright fisher con una poblacion de 3 gener, encuentre la probabilidad que no no halla A llletes en 3 generaciones


Veamos como crear la matriz

```{r}
factorial_funcion<-function(k,i,j){

  a <- (factorial(k)/factorial(j))*factorial(k-j)
  b <- (i/k)^j
  c <- (1 - (i/k))^(k-j)
  
  print(a*b*c)
} 

```


```{r}
#solo los estados correspondientes
aa<-(factorial_funcion(3,1,2))
ab<-(factorial_funcion(3,2,1))
ac<-(factorial_funcion(3,2,2))
ad<-(factorial_funcion(3,2,3))
ae<-(factorial_funcion(3,1,3))
 
matriz2 <- matrix(c(1,ad,ae,0,0,ac,aa,0,0,aa,ac,0,0,ae,ad,1),  nrow = 4 ,ncol = 4,)
matriz2
```
b)
```{r}
matriz2_cuadrada <- matriz2%*%matriz2%*%matriz2
matriz2_cuadrada
#localizar 3,1
```
Observando veemos que no exista A (alelo) es de 0.516 o 51.6%

Example 2.18 (Lung cancer study)
Medical researchers can use simulation to
study the progression of lung cancer in the body, as described in Example 2.12. The
50 × 50 transition matrix is stored in an Excel spreadsheet and can be downloaded
into R from the ile cancerstudy.R. The initial distribution is a vector of all 0s with
a 1 at position 23, corresponding to the lung. See the documentation in the script ile
for the 50-site numbering system. Common sites are 24 and 25 (lymph nodes) and
22 (liver). Following are several simulations of the process for eight steps.


```{r}
mat <- read.csv("lungcancer.csv",header=T)
init <- c(rep(0,22),1,rep(0,27)) # Starting state 23 is Lung
n <- 8
markov(init,mat,n)
```
Graduados ->
University administrators have developed a Markov model to simu-
late graduation rates at their school. Students might drop out, repeat a year, or move
on to the next year. Students have a 3% chance of repeating the year. First-years and
sophomores have a 6% chance of dropping out. For juniors and seniors, the drop-out
rate is 4%. 

```{r}
init <- c(0,1,0,0,0,0) #Student starts as first-year
P <- matrix(c(1,0,0,0,0,0,0.06,0.03,0.91,0,
0,0,0.06,0,0.03,0.91,0,0,0.04,0,0,0.03,0.93,0,
0.04,0,0,0,0.03,0.93,0,0,0,0,0,1),nrow=6,byrow=T)
states <-
c("Drop","Fr","So","Jr","Se","Grad")
 rownames(P) <- states
 colnames(P) <- states

markov(init,P,10,states)

sim <- replicate(10000,markov(init,P,10,states)[11])
table(sim)/10000


```


Simulacion de gamblers Ruin ->

```{r}
gen.ruin = function(n, x.cnt, y.cnt, x.p){
  x.cnt.c = x.cnt
  y.cnt.c = y.cnt
  x.rnd = rbinom(n, 1, p=x.p)
  x.rnd[x.rnd==0] = -1
  y.rnd = x.rnd*-1
  x.cum.sum = cumsum(x.rnd)+x.cnt
  y.cum.sum = cumsum(y.rnd)+y.cnt
  
  ruin.data = cumsum(x.rnd)+x.cnt
  
  if( any( which(ruin.data>=x.cnt+y.cnt)) | any(which(ruin.data<=0))){cut.data = 1+min( which(ruin.data>=x.cnt+y.cnt), which(ruin.data<=0) )
  
  ruin.data[cut.data:length(ruin.data)] = 0
  
  }
  
  return(ruin.data)
  
  }
n.reps = 10000
ruin.sim = replicate(n.reps, gen.ruin(n=1000, x.cnt=5, y.cnt=10, x.p=.6))
ruin.sim[ruin.sim==0] = NA
hist( apply(ruin.sim==15 | is.na(ruin.sim), 2, which.max) , nclass=100, col='8', main="Distribution of Number of Turns",
xlab="Turn Number")
abline(v=mean(apply(ruin.sim==15 | is.na(ruin.sim), 2, which.max)), lwd=3, col='red')
abline(v=median(apply(ruin.sim==15 | is.na(ruin.sim), 2, which.max)), lwd=3, col='green')
x.annihilation = apply(ruin.sim==15, 2, which.max)
( prob.x.annilate = length(x.annihilation[x.annihilation!=1]) / n.reps )
state.cnt = ruin.sim
state.cnt[state.cnt!=15] = 0
state.cnt[state.cnt==15] = 1
mean.state = apply(ruin.sim, 1, mean, na.rm=T)
plot(mean.state, xlim=c(0,which.max(mean.state)), ylim=c(0,20), ylab="Points", xlab="Number of Plays", pch=16, cex=.5, col='green')
lines(mean.state, col='green')
points(15-mean.state, pch=16, cex=.5, col='blue')
lines(15-mean.state, col='blue')
```

Changes in the distribution of wetlands in Yinchuan Plain, China are studied in Zhang et al. (2011). Wetlands are considered among the most important ecosystems on earth. A Markov model is developed to track yearly changes in wetland type. Based on imaging and satellite data from 1991, 1999, and 2006, researchers measured annual distributions of wetland type throughout the region and estimated the Markov transition matrix

The state Non refers to nonwetland regions. Based
dict that “The wetland distribution will essentially be in a Plain in approximately 100 years.”
0.005 0.001 0.020 0.252 0.107 0.005 0.043 0.508 0.015 0.002 0.004 0.665 0.007 0.007 0.025
With technology one checks that P100 = P101 has identical rows. The common row gives the predicted long-term, steady-state wetland distribution


```{r}
P <- matrix(c(0.342,0.005,0.001,0.020,0.632,0.001,0.252,0.107,0.005,0.635,0,0.043,0.508,0.015,0.434,0.001,0.002,0.004,0.665,0.328,0.007,0.007,0.007,0.025,0.954),nrow=5,byrow=T)
P
```

```{r}
matrixpower(P,100)
```

1.33 Cards are drawn from a standard deck,with replacement,until an ace appears. Simulate the mean and variance of the number of cards required.



Numero de cartas = 52

 => n(s) = 52
 Dejemos E un evento de terner un as de proabilidad
 
 = $n(E)/n(s) = 4/52 = 1/13$
  
Probabilidad de no tener un as

= 1 - $1/13$ = $12/13$

Sabemos que n = 2

Media = np = 2($1/13$) = $2/13$
Varianza = npq = 2($1/13$)($12/13$)

1.37 Solve 1.28
On any day, the number of accidents on the highway has a Poisson distribution with parameter Λ. The parameter Λ varies from day to day and is itself a random variable. Find the mean and variance of the number of accidents per day when Λ is uniformly distributed on (0, 3


$X = numero de accidentes / dia$ 

\[
\sum_{k=0}^{2}x_{i} = 1 - (x(-^3)*3^k)/k! = 1 - 0.57681 = 0.42319
\]

$ e^-3 * 3 ^ 0 / 0! = 0.0497$
$ e^-3 * 3 ^ 1 / 1! = 0.1493$
$ e^-3 * 3 ^ 2 / 2! = 0.2240$
$ e^-3 * 3 ^ 3 / 3! = 0.2240$

```{r}
#rango de poisson 
rango <- 0:3
#crear una grafica
plot(rango, dpois(rango, lambda=4), type='h')
```
```{r}
poissonone <-c( dpois(0, 1) , dpois(1, 2) , dpois(2, 3) ,  dpois(3, 4))
mediana = median(poissonone)
varianza = var(poissonone)

sprintf("Mediana %f" , mediana)
sprintf("Varianza %f" , varianza)
```

2.25 .- the behavior of dolphins in the presence of tour boats in Patagonia, Argentina is studied in Dans et al. (2012). A Markov chain model is devel- oped, with state space consisting of five primary dolphin activities (socializing,
EXERCISES 75 traveling, milling, feeding, and resting). The following transition matrix is
obtained.

2.27 ->

R : See gamblersruin.R. Simulate gambler’s ruin for a gambler with initial stake $2, playing a fair game.
(a) Estimate the probability that the gambler is ruined before he wins $5.
(b) Construct the transition matrix for the associated Markov chain. Estimate
the desired probability in (a) by taking high matrix powers. (c) Compare your results with the exact probability.

```{r}
# gamble(k, n, p)
  #   k: Gambler's initial state
  #   n: Gambler plays until either $n or Ruin
  #   p: Probability of winning $1 at each play
  #   Function returns 1 if gambler is eventually ruined
  #                    returns 0 if gambler eventually wins $n
  
gamble <- function(k,n,p) {
        stake <- k
        while (stake > 0 & stake < n) {
                bet <- sample(c(-1,1),1,prob=c(1-p,p))
                stake <- stake + bet
        }
        if (stake == 0) return(1) else return(0)
        }   
```

```{r}
k <- 2

n <- 5  

p <- 1/2  

trials <- 1000

simlist <- replicate(trials, gamble(k, n, p))

mean(simlist) # Estimate of probability that gambler is ruined
```
Using Gamblers ruin probability will be can be calculated by taking higher powers of the transition prob matrix and value taken transition from 2 to 0. The code is given below
```{r}
tpm <- matrix(c(1, 0, 0, 0, 0, 0,

+ 0.5, 0, 0.5, 0 ,0 ,0,

+ 0, 0.5, 0 ,0.5, 0, 0,

+ 0, 0, 0.5, 0, 0.5, 0,

+ 0, 0, 0, 0.5, 0, 0.5,

+ 0, 0, 0, 0, 0, 1),byrow = T,nrow = 6)
```

```{r}
 library(expm)   
  p5<- tpm %^% 5
  p5
```


```{r}
p15<- tpm %^% 15

p15
```

```{r}
p25<- tpm %^% 25

p25
```

```{r}
p50<- tpm %^% 50
p50
```

```{r}
p5[3,1]
```
```{r}
p15[3,1]
```

```{r}
 p25[3,1]
```

```{r}
p50[3,1]
```

De los resultados obtenenidos los resultados de probabilidad convergen en 0.6 como incrementos

La probabilidad exacta es (n-k)/n=(5-2)/5=0.6

Resumen

Exact0 = 0.6

Simulacion =0.59

Usando Markov Chain, tpm^5 = 0.375, tpm^15=.573, tpm^25=0.596, tpm^50=0.599